\chapter{Literature Review}

In the developmental history of stochastic volatility models, the 3/2 model has increasingly become a focal point of academic research due to its unique non-linear characteristics and robust capacity to fit extreme market conditions. This chapter systematically reviews the domestic and international literature concerning the 3/2 model and related option pricing methodologies, first outlining the development of the model and its extended frameworks, and subsequently exploring the evolution of various exact simulation methods.

\section{Development of the 3/2 Model and its Extended Frameworks}

To address the limitations of traditional volatility models in characterizing complex market dynamics, numerous scholars have dedicated efforts to improving and extending the 3/2 model from the perspectives of model structure and analytical frameworks.

* Alternative Modeling and Asymptotic Expansion: Carr and Sun (2007) developed a novel framework that bypasses the need to directly model instantaneous volatility or specify the market price of volatility risk by modeling the variance swap rate directly. Concurrently, Medvedev and Scaillet (2007) proposed a comprehensive framework based on asymptotic expansion to address the option pricing challenges inherent in non-affine local stochastic volatility models, including the 3/2 model.
* Introduction of Jumps and Unified Models: To achieve consistent modeling of equity and VIX derivatives, Baldeaux and Badran (2012) introduced the "3/2 plus jumps model". More representatively, Grasselli (2017) proposed the "4/2 stochastic volatility model", a versatile framework that subsumes both the Heston and the 3/2 models as special cases. By assuming the instantaneous variance is a linear combination of the 1/2 and 3/2 terms, this "4/2 model" effectively leverages the strengths of both predecessors.
* Handling Non-affine Characteristics: Addressing the non-affine property of the 3/2 model, Kouarfate, Kouritzin, and Mackay (2021) developed a novel simulation framework based on an explicit weak solution for option pricing. This framework ingeniously leverages the property that the variance process of the 3/2 model is the inverse of a CIR process.

\section{Evolution of Exact Simulation Methods}

Having established the foundational framework of the 3/2 model, achieving efficient pricing while avoiding time discretization errors became another major research priority. Academia has subsequently developed and continuously optimized a series of exact simulation methods.

* Establishment of the Foundational Framework: Broadie and Kaya (2006) suggested a method for the exact simulation of the stock price and variance under Heston’s stochastic volatility model and other affine jump diffusion processes, which is known as the exact MC scheme.
* Initial Efficiency Improvements: Subsequently, Glasserman and Kim (2011) significantly improved simulation efficiency by avoiding the numerical inverse Laplace transform. Using results derived by Pitman and Yor (1982), they decomposed the conditional integrated variance into an infinite series of gamma random variables.
* Extension and Application to the 3/2 Model: Targeting the 3/2 model specifically, Baldeaux (2012) discussed the exact simulation of the stock price process underlying the model. Using a result derived by Craddock and Lennox via Lie Symmetry Analysis, he successfully adapted the Broadie-Kaya algorithm for affine processes to the 3/2 model.
* Recent Breakthroughs and Innovative Algorithms: In recent years, algorithmic optimization in this field has achieved further breakthroughs. Choi J and Kwok YK (2024) proposed a new exact simulation scheme without utilizing the modified Bessel function, based on the core observation that the conditional integrated variance can be simplified when conditioned by the Poisson variate used for simulating the terminal variance. Furthermore, Brignone and Junike (2026) proposed a novel simulation framework leveraging the conditional COS method. This research addresses long-standing implementation challenges in exact simulation, particularly concerning error control and computational overhead observed since Broadie and Kaya (2006).

\section{Current State of Domestic Research}

Compared to the abundant achievements internationally, domestic research on stochastic volatility models started relatively later, although significant progress has been made in recent years. However, the majority of current domestic studies focus on the Heston and SABR models, with very few literatures deeply exploring the 3/2 volatility model. This current state indicates a substantial gap in the research of the 3/2 model and its efficient pricing algorithms within the domestic financial engineering field, providing broad scope and practical significance for this thesis.

\section{Literature Review Summary and Research Positioning}

Synthesizing the aforementioned literature, it is evident that despite the fruitful results achieved by the international academic community in extending the theoretical framework of the 3/2 volatility model (e.g., introducing jump processes, constructing the unified 4/2 model) and optimizing exact simulation algorithms (e.g., avoiding inverse Laplace transforms, introducing the conditional COS method), the existing option pricing systems still possess distinct limitations. Specifically, the five mainstream pricing methods frequently encounter bottlenecks—such as computational failure, uncontrollable errors, or severe efficiency degradation—when faced with extreme market scenarios (e.g., extremely short times to expiration, ATM option pricing, and specific extreme parameter combinations). Meanwhile, systematic research in this domain remains scarce domestically, presenting an urgent need to introduce and improve cutting-edge pricing technologies.

Based on these research gaps, the positioning of this study is to: break through the applicability barriers of existing pricing methods in specific extreme scenarios, find the optimal solution balancing computational accuracy and efficiency, and propose a highly robust and fast pricing mechanism.

Distinguished from previous studies that merely stopped at superficial algorithmic improvements, this research delves into the underlying mathematical logic, specifically making the following core contributions:

\begin{enumerate}
    \item Analysis of Underlying Mathematical Properties: Systematically analyzing the first-order partial and second-order derivative of the Modified Bessel Function of the first kind for the first time, deeply studying its mathematical properties, and deriving explicit convergence conditions.
    \item Proposal of an All-Scenario Fast Pricing Framework: Based on the aforementioned theoretical breakthroughs, presenting a novel, fast option pricing method that perfectly covers various extreme parameter scenarios under the 3/2 volatility model, resolving the failures of previous methods in pricing ATM and near-expiration options.
    \item Multi-dimensional Comprehensive Benchmarking: Through extensive numerical experiments, comprehensively comparing the performance of mainstream pricing methods across different scenarios, clearly revealing the trade-offs between computational accuracy and efficiency among different algorithms, thereby providing a solid empirical basis for selecting the optimal pricing strategy in diverse application contexts.
\end{enumerate}