\chapter{Contributions and Conclusion}
\label{chap:conclusion}

\section{Core Innovations and Contributions}

Moving beyond a purely empirical comparison of existing algorithms, this thesis fundamentally resolves several mathematical bottlenecks inherent in the $3/2$ stochastic volatility framework. The primary theoretical and methodological advancements of this research are tripartite:

\begin{enumerate}
    \item Deconstruction of the Modified Bessel Function: A chronic hurdle in exact simulation literature is the intractability of complex special functions. This study breaks new ground by rigorously mapping the first-order partial and second-order derivatives of the Modified Bessel Function of the first kind.
    \item Formalization of Convergence Thresholds: Grounded in the analytical deconstruction of the Bessel function, we successfully derive and identify explicit mathematical convergence conditions. This theoretical scaffolding is critical, providing the strict mathematical guarantee required to stabilize subsequent simulation architectures.
    \item A Universal and Accelerated Valuation Framework: Synthesizing these mathematical insights, we introduce an unconditionally robust option pricing mechanism. By neutralizing the structural defects that plague conventional exact simulations and standard FFT integrations, this proposed engine guarantees reliable, high-speed valuations across the entire parameter spectrum of the $3/2$ model, including the most extreme market regimes.
\end{enumerate}

\section{Empirical Findings and Research Conclusions}

Our empirical campaign, built upon seven distinct stress-test matrices (Cases I through VII), critically benchmarks five prevailing numerical techniques against our enhanced methodology. The resulting data explicitly exposes the stark operational realities and trade-offs of current pricing engines:

\begin{itemize}
    \item Generic Euler/Milstein schemes incur an unacceptable computational drag, rendering them obsolete for latency-sensitive applications.
    \item Traditional exact and "almost exact" simulation routines routinely collapse—yielding severe biases or outright failing—when forced to price at-the-money (ATM) or short-dated derivatives.
    \item Similarly, while the pure FFT boasts near-instantaneous execution under normal conditions, it fundamentally disintegrates near maturity due to uncontrollable oscillatory integration failures.
\end{itemize}

In conclusion, the empirical evidence dictates that our proposed conditional Fourier-Cosine strategy successfully reconciles the chronic tension between pricing fidelity and execution speed. By mapping out these algorithmic boundaries and validating a mathematically sound alternative, this research pushes the theoretical frontier of stochastic volatility modeling. More importantly, it equips quantitative desks with a highly dependable valuation toolkit, directly applicable to demanding practical environments such as high-frequency market making and real-time portfolio risk management.