\chapter{The 3/2 Volatility Model Specification}
\label{chap:model}

\section{Evolution and Economic Intuition of the 3/2 Model}
\label{sec:model_setup}

As an advanced stochastic volatility framework, the 3/2 model emerged as a sophisticated evolution of the Square-Root Process (popularized by Cox-Ingersoll-Ross and later further developed by Steven Heston).

In the current financial industry, the Heston model (a 1/2 model) has long been regarded as the industry standard. However, during periods of extreme market stress, the Heston model often fails to accurately capture the "volatility of volatility" observed in the market and struggles to perfectly fit the steepness of the short-term volatility skew.

To overcome this limitation, researchers in the late 1990s and early 2000s proposed a novel approach: shifting the exponent of the diffusion term from 1/2 to 3/2. This modification in specification allows the model to generate more explosive volatility dynamics and a non-linear mean-reversion speed. Consequently, the 3/2 model allows for the occurrence of sharp peaks in volatility and presents a steeper skew under market pressure, which is highly consistent with empirical data. However, it is worth noting that under certain normal market conditions, the 3/2 model may lack the stability inherent in the Heston model.

\section{Stochastic Differential Equation (SDE) Specifications}

Under the risk-neutral measure, the asset price process and the instantaneous variance/volatility process of the 3/2 model can be characterized by the following system of Stochastic Differential Equations (SDEs):The dynamic evolution of the underlying asset price process $S_t$ satisfies:

$$dS_t = r S_t dt + S_t \sqrt{V_t} (\rho dW_t^1 + \sqrt{1-\rho^2} dW_t^2)$$

where $r$ represents the risk-free interest rate, and $\rho$ is the correlation coefficient between the two Brownian motions.

Simultaneously, the dynamic evolution of the variance process $V_t$ satisfies:

$$dV_t = \kappa V_t(\theta - V_t)dt + \epsilon V_t^{3/2}dW_t^1$$

In the above equations:

\begin{itemize}
    \item $\kappa$ controls the speed of mean reversion.
    \item $\theta$ represents the long-term average variance level.
    \item $\epsilon$ (or denoted as $\nu$) is the volatility of volatility (vov) parameter, controlling the fluctuation degree of the variance process.
    \item $dW_t^1$ and $dW_t^2$ are independent standard Brownian motions.
\end{itemize}

It is worth noting that this variance process can also be equivalently expressed in terms of $dV_t/V_t$, thereby more intuitively reflecting the drift and diffusion characteristics of its relative changes.

\section{Integrated Variance and Conditional Distribution}

In the option pricing framework of the 3/2 model, the integrated variance $V_T$ over the time interval $[0, T]$ plays a crucial role. It is defined as follows:

$$V_T = \int_0^T \sigma_t^2 dt = \int_0^T V_t dt$$

Since the randomness of the asset price is partially driven by volatility, when we condition on the integrated variance $V_T$, the terminal price of the underlying asset $S_T$ follows a lognormal distribution. This pivotal property implies that, conditional on $V_T$, we can leverage the pricing framework of the Black-Scholes (BS) model.

According to Itô's isometry principle:

$$\int_0^T \rho^* \sigma_t dX_t \sim \rho^* \sqrt{V_T} X_1 \sim N(0, (\rho^*)^2 V_T)$$

Accordingly, the effective volatility $\sigma$ between time $0$ and $T$ can be expressed as:

$$\sigma = \rho^* \sqrt{V_T / T}$$

Through further stochastic calculus derivation, the formula for the log return of the terminal price relative to the initial price can be expanded as:

$$\log\left(\frac{S_T}{S_0}\right) = \frac{\rho}{\epsilon}\log\left(\frac{V_T}{V_0}\right) + \left(\kappa + \frac{\epsilon^2}{2}\right)V_T - \kappa\theta T - \frac{1}{2}V_T + \rho^* \sqrt{V_T} X_1$$

This formula serves as the core foundation for subsequent exact simulation and Fourier transform pricing.