\chapter{Introduction}

\section{Background and Significance}

The evolution of financial engineering has long positioned option pricing as its central pillar. Historically, the Black-Scholes (BS) framework dominated this landscape, primarily valued for its elegant, closed-form solutions. However, the model’s reliance on the assumption of constant volatility represents a significant departure from real-world market behavior. Extensive empirical observations by both academics and market participants reveal that asset volatility is far from static; instead, it manifests as a complex stochastic process. This inherent randomness introduces substantial layers of difficulty when attempting to model modern financial markets accurately.

To rectify the shortcomings of the constant volatility premise, a variety of stochastic volatility frameworks have been introduced. Among these, the 3/2 volatility model has emerged as a particularly compelling subject of study. 
Distinguishing itself from traditional 1/2 models—such as the Heston model—the 3/2 framework posits that volatility dynamics follow an inverse 3/2 power law. This specific non-linear architecture provides the model with a superior capacity to track explosive volatility spikes during periods of extreme market turbulence. Consequently, it offers a more robust fit for the \enquote{volatility smile} observed in empirical data compared to its affine counterparts.

\section{Pricing Challenges under the 3/2 Volatility Model}

While the 3/2 model offers superior performance in representing market dynamics, its sophisticated mathematical architecture imposes significant computational burdens on derivative valuation. Currently, the academic and practitioner communities rely on five primary methodologies for option pricing within this framework, yet each encounters specific bottlenecks that hinder their universal application:

\begin{enumerate}
    \item Euler/Milstein scheme: As a generic numerical approach, these schemes suffer from notoriously slow convergence rates, making them computationally prohibitive for real-time risk management.
    \item Exact simulation: Although designed to eliminate time-discretization errors, this method is plagued by sluggish execution speeds. More critically, it exhibits structural instability under specific parameter sets, often failing to produce valid prices for at-the-money (ATM) options or those nearing maturity.
    \item Almost exact simulation: This hybrid approach, intended to refine the simulation process, frequently collapses under specific parameter combinations, rendering it incapable of completing the pricing task.
    \item Exact simulation based on conditional Fourier-cosine Method: This hybrid approach, intended to refine the simulation process, frequently collapses under specific parameter combinations, rendering it incapable of completing the pricing task.
    \item Fast Fourier Transform (FFT): Despite its reputation for high-speed calculation, the FFT method possesses a fundamental vulnerability: the oscillatory nature of the integrand causes massive numerical inaccuracies or outright failure when pricing options with very short times to expiration.
\end{enumerate}

\section{Research Objective and Main Content}

Motivated by the operational failures and severe latencies of standard pricing algorithms under boundary conditions, this thesis seeks to overhaul the valuation framework for the $3/2$ stochastic volatility model.

Our research trajectory is anchored in two primary phases. First, we reconstruct the five prevailing numerical schemes, benchmarking their performance against vanilla options across a diverse grid of market regimes. By dissecting the resulting simulation data, the study maps out the exact compromise each algorithm makes between computational throughput and pricing fidelity, providing a transparent view of their operational boundaries.

The core contribution of this work, however, is a methodological breakthrough designed specifically to bypass the severe limitations these traditional models face with at-the-money (ATM) and short-tenor options. To resolve these bottlenecks, we introduce a highly robust pricing mechanism. This advanced approach consistently delivers rapid, high-precision valuations, retaining its structural integrity regardless of whether the market environment is tranquil or under extreme stress.