\chapter{Pricing Methodologies}
\label{chap:analysis}

With the fundamental stochastic differential equations and the integrated variance properties of the $3/2$ volatility model now established, the focus of this study shifts to practical numerical implementation. Here, we translate these theoretical constructs into actionable option pricing algorithms. The following sections outline the precise mathematical derivations and computational steps for two distinct valuation paradigms: time-domain Monte Carlo architectures—spanning both conditional and exact simulation schemes—and frequency-domain techniques driven by Fourier transformations.

\section{Monte Carlo Simulation Procedures}
\label{sec:agent_decision_analysis}

Under the 3/2 model, directly discretizing the original SDEs (e.g., using the Euler scheme) often introduces significant truncation errors. Therefore, this study primarily investigates two advanced schemes: conditional Monte Carlo simulation and exact Monte Carlo simulation.

\subsection{Conditional MC}
Conditional Monte Carlo simulation aims to approximate the integrated variance by generating volatility paths. Its primary execution steps are as follows:
\begin{enumerate}
  \item Path Simulation: Simulate the discrete paths of the volatility process $\sigma_t$ over the time grid $0 \le t_1 \le t_2 \le \dots \le T$, and obtain the terminal volatility $\sigma_T$.
  \item Integral Calculation: Utilize numerical integration rules (such as the Trapezoidal or Simpson’s rule) to perform time integration on the simulated variance paths, thereby obtaining the integrated variance $V_T$ over the entire time period.
\end{enumerate}

\subsection{Exact MC}
To completely eliminate time discretization errors, the exact simulation method directly samples the terminal states without relying on the step-by-step generation of paths. Its core steps are:
\begin{enumerate}
  \item Terminal Variance Simulation: Directly simulate the terminal variance $V_T$ using the known transition density of the variance process under the 3/2 model.
  \item Conditional Integrated Variance Sampling: Conditional on the known $v_T$ (or equivalently, the terminal volatility $\sigma_T$), sample the integrated variance $V_T$ from the numerical cumulative distribution function (CDF) of $V_T|v_T$. This step typically involves complex inner-layer Inverse Laplace transform operations.
\end{enumerate}

\subsection{Common Procedures for Pricing and Aggregation}
Regardless of whether the integrated variance is obtained via conditional MC or exact MC, the subsequent steps for calculating the option price are universal:
\begin{enumerate}[start=3]
  \item Parameter Transformation: Calculate the expected value of the underlying asset conditional on given $v_T$ and $V_T$, denoted as $E(S_T|v_T, V_T)$, and determine the effective volatility, usually expressed as $\rho^*\sqrt{V_T/T}$.
  \item Option Valuation:
  \begin{itemize}
    \item Approach 1 (Price Sampling): Draw samples from the lognormal or normal distribution to calculate the terminal option payoff.
    \item Approach 2 (Analytical Formula): Directly substitute the modified parameters into the Bachelier or Black-Scholes-Merton (BSM) option pricing formulas. Specifically, set the modified initial price as $S_0 := E(S_T|v_T)$ and the modified volatility as $\sigma := \rho^*\sqrt{V_T/T}$.
  \end{itemize}
  \item Averaging: Average the resulting option prices across all independent simulation paths to obtain the final expected present value of the option.
\end{enumerate}

\section{Fourier Transformation Procedures}
Assuming the characteristic function is accessible, frequency-domain techniques offer a compelling, high-efficiency alternative to traditional time-domain Monte Carlo simulations.

\begin{enumerate}
    \item Fast Fourier Transform (FFT): The primary advantage of the FFT lies in its ability to evaluate prices across a spectrum of strikes simultaneously by mapping the valuation problem into Fourier space. Yet, this computational speed comes at a steep cost for short-dated options. As maturity approaches zero, the integrand becomes highly oscillatory, triggering massive numerical instability and effectively breaking the valuation mechanism.
    \item Fourier-Cosine Transformation (COS Method): Because the standard FFT collapses under short maturities, our research deploys the conditional Fourier-Cosine expansion as a structural remedy. By reconstructing the underlying probability density function via a truncated cosine series, this technique inherently bypasses the instability caused by highly oscillatory integrals. As our subsequent numerical experiments will confirm, the COS method strikes an optimal balance—retaining the rapid execution typical of frequency-domain approaches while delivering robust, high-precision valuations even at the absolute boundaries of the parameter space.
\end{enumerate}